Population Synthesis Literature Review

### 2017-12-15 review/review.tex
added review documents to tree

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\graphicspath{{./images/}}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{listings}
\usepackage[english]{babel}
\usepackage{enumerate}
\lstset{frame=tlrb,xleftmargin=\fboxsep,xrightmargin=-\fboxsep}

\usepackage[style=numeric, backend=biber, maxcitenames=2]{biblatex}
\bibliography{bib/references}

\begin{document}

@@

The population synthesis method that \textcite{petrik2016measuring} used adapted ideas from probabilistic graphical models (PGM).
In their graphical model, there are 7 socio-demographic attributes, which are Age, Gender, Employment, Income, Car ownership, Purpose of trip, and Who pay for the trip, to represent a person.
Once the dependencies between variables are specified, we need to calculate or estimate conditional probability of one node given its parents.
The flexibility of this method is that those conditional probabilities can be derived from different sources like census data or surveys on individuals.
Finally, the synthetic population is just random draws from the probabilistic graphical model.
Though we can take marginal distributions of all variables as supplement for the graphical model, the population generated by Bayesian network model can only precisely match the marginals of mother nodes.
This is because the distribution of a node is uniquely defined by its parent node and the conditional distribution based on PGM.
And this point should be taken into consideration when choosing the population synthesis method.

\section{Data}
Most paper in population synthesis used the Public Use Microdata Sample (PUMS), which is available both in individual level and household level.
It includes multiple socio-economic parameters of a household or an individual in the region.
The sample rate of this dataset is about 1\% to 10\% of the total population in the region.
We found PUMS data is complete in US cities but haven't found microdata in other cities in the world.
Currently, we have found aggregated information of population, age, sex, household, education in districts in Beijing.

The choice of algorithm should be based both on the data we have and the appearance of the final result we want.
What is also important for us is the simplicity and novelty of methodology.

@@

In the next step, the synthetic population shall be constructed by selecting households from PUMS in proportion to the estimated weights in the multiway table.
Specifically, the number of households of each demographic type in a census tract can be obtained by multiplying the total number of households by the probability in the cell representing this household type, or by drawing the numbers at random according to these probabilities.
However, the household type represent by a cell in multiway table is not exactly that of PUMS data.
The fitted multiway table can only be span by control variables, which is defined as variables which have marginal distribution in census data.
However, households and individuals in disaggregate data are usually described by additional desired variables.
So, the probability of a household type in PUMS data being chosen shall be assigned according to the distance between such a household $p$ and a household type $c$ in multi-way table.
In \cite{beckman1996creating}, this probability is defined as 
\begin{equation}
	D(p,c) = w_p \prod_{i \in J}\Big(1-|(d_i^p-d_i^c)/r_i|^k\Big)*\prod_{i \notin J}\Big(1-\Delta(d_i^p, d_i^c)\Big)
\end{equation}

@@

\begin{enumerate}[Step 1:]
\item Network: modify OSM network file, make it legal input to SimMobility
  \begin{enumerate}[1.]
  \item links has to be unidirectional
  \item shorten the links and add turning points
  \item maintain the attributes of each object in the network
  \end{enumerate}
\item Population Synthesis
  \begin{enumerate}[1. ]
  \item Data collection: survey data, aggregate census data for zones in each city
  \item Learn the parameter and the structure of Bayesian Network which captures the structure of population
  \item sampling from Bayesian Network to get whole population in the city
  \item Clustering the synthetic population in a city by K-means
  \item Divide the city into zones according to clustering result, and form the aggregate census data for zones
  \end{enumerate} 
\end{enumerate}

@@

@@

\newpage
\section{References}
\printbibliography[heading=none]
\end{document}

>>>>

\documentclass[11pt,twoside]{article}
\usepackage{etex}
\newcommand{\num}{6{} }

\raggedbottom

%geometry (sets margin) and other useful packages
\usepackage{geometry}
\geometry{top=1in, left=1in,right=1in,bottom=1in}
 \usepackage{graphicx,booktabs,calc}

%=== GRAPHICS PATH ===========
\graphicspath{{./images/}}
% Marginpar width
%Marginpar width
\newcommand{\pts}[1]{\marginpar{ \small\hspace{0pt} \textit{[#1]} } } 
\setlength{\marginparwidth}{.5in}
%\reversemarginpar
%\setlength{\marginparsep}{.02in}

%% Fonts
% \usepackage{fourier}
% \usepackage[T1]{pbsi}

\usepackage{lmodern}
\usepackage[T1]{fontenc}
%\usepackage{minted}

\usepackage{rotating}
%% Cite Title
\usepackage[style=authoryear,maxcitenames=2,natbib]{biblatex}
\addbibresource{bib/references.bib}

%%% Counters
\usepackage{chngcntr,mathtools}
\counterwithout{figure}{section}
\counterwithout{table}{section}

\numberwithin{equation}{section}

%% Captions
\usepackage{caption}
\captionsetup{
  labelsep=quad,
  justification=raggedright,
  labelfont=sc
}

%AMS-TeX packages
\usepackage{amssymb,amsmath,amsthm} 
\usepackage{bm}
\usepackage[mathscr]{eucal}
\usepackage{colortbl}
\usepackage{color}

\usepackage{epstopdf,subfigure,hyperref,enumerate,polynom,polynomial}
\usepackage{multirow,minitoc,fancybox,array,multicol}

\definecolor{slblue}{rgb}{0,.3,.62}%
%\renewcommand{\chaptermark}[1]{ \markboth{#1}{} }
\renewcommand{\sectionmark}[1]{ \markright{#1}{} }

\usepackage{fancyhdr}
\pagestyle{fancy}
%\addtolength{\headwidth}{\marginparsep} %these change header-rule width
%\addtolength{\headwidth}{\marginparwidth}
%\fancyheadoffset{30pt}
%\fancyfootoffset{30pt}
\fancyhead[LO,RE]{\small  \it \nouppercase{\leftmark}}
\fancyhead[RO,LE]{\small Page \thepage} 
\fancyfoot[RO,LE]{\small }% PR \num S-2015} 
\fancyfoot[LO,RE]{\small }%\scshape MODL} 
\cfoot{} 
\renewcommand{\headrulewidth}{0.1pt} 
\renewcommand{\footrulewidth}{0pt}
%\setlength\voffset{-0.25in}
%\setlength\textheight{648pt}

\usepackage{paralist}

%%% FORMAT PYTHON CODE
\usepackage{listings}
% Default fixed font does not support bold face
\DeclareFixedFont{\ttb}{T1}{txtt}{bx}{n}{8} % for bold
\DeclareFixedFont{\ttm}{T1}{txtt}{m}{n}{8}  % for normal

% Custom colors
\usepackage{color}
\definecolor{deepblue}{rgb}{0,0,0.5}
\definecolor{deepred}{rgb}{0.6,0,0}
\definecolor{deepgreen}{rgb}{0,0.5,0}

%\usepackage{listings}

% % Python style for highlighting
% \newcommand\pythonstyle{\lstset{
% language=Python,
% basicstyle=\footnotesize\ttm,
% otherkeywords={self},             % Add keywords here
% keywordstyle=\footnotesize\ttb\color{deepblue},
% emph={MyClass,__init__},          % Custom highlighting
% emphstyle=\footnotesize\ttb\color{deepred},    % Custom highlighting style
% stringstyle=\color{deepgreen},
% frame=tb,                         % Any extra options here
% showstringspaces=false            % 
% }}

% % Python environment
% \lstnewenvironment{python}[1][]
% {
% \pythonstyle
% \lstset{#1}
% }
% {}

% % Python for external files
% \newcommand\pythonexternal[2][]{{
% \pythonstyle
% \lstinputlisting[#1]{#2}}}

% % Python for inline
% \newcommand\pythoninline[1]{{\pythonstyle\lstinline!#1!}}

\newcommand{\osn}{\oldstylenums}
\newcommand{\dg}{^{\circ}}
\newcommand{\lt}{\left}
\newcommand{\rt}{\right}
\newcommand{\pt}{\phantom}
\newcommand{\tf}{\therefore}
\newcommand{\?}{\stackrel{?}{=}}
\newcommand{\fr}{\frac}
\newcommand{\dfr}{\dfrac}
\newcommand{\ul}{\underline}
\newcommand{\tn}{\tabularnewline}
\newcommand{\nl}{\newline}
\newcommand\relph[1]{\mathrel{\phantom{#1}}}
\newcommand{\cm}{\checkmark}
\newcommand{\ol}{\overline}
\newcommand{\rd}{\color{red}}
\newcommand{\bl}{\color{blue}}
\newcommand{\pl}{\color{purple}}
\newcommand{\og}{\color{orange!90!black}}
\newcommand{\gr}{\color{green!40!black}}
\newcommand{\nin}{\noindent}
\newcommand{\la}{\lambda}
\renewcommand{\th}{\theta}
\newcommand{\al}{\alpha}
\newcommand{\G}{\Gamma}
\newcommand*\circled[1]{\tikz[baseline=(char.base)]{
            \node[shape=circle,draw,thick,inner sep=1pt] (char) {\small #1};}}

\newcommand{\bc}{\begin{compactenum}[\quad--]}
\newcommand{\ec}{\end{compactenum}}

\newcommand{\p}{\partial}
\newcommand{\pd}[2]{\frac{\partial{#1}}{\partial{#2}}}
\newcommand{\dpd}[2]{\dfrac{\partial{#1}}{\partial{#2}}}
\newcommand{\pdd}[2]{\frac{\partial^2{#1}}{\partial{#2}^2}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

@@

The population synthesis method that %\cite{petrik2016measuring}
used adapted ideas from probabilistic graphical models (PGM).
In their graphical model, there are 7 socio-demographic attributes, which are Age, Gender, Employment, Income, Car ownership, Purpose of trip, and Who pay for the trip, to represent a person.
Once the dependencies between variables are specified, we need to calculate or estimate conditional probability of one node given its parents.
The flexibility of this method is that those conditional probabilities can be derived from different sources like census data or surveys on individuals. Nevertheless, compared to the high dependency of IPF on the quality of sample data, Bayesian method gives synthetic population more heterogeneity.
Finally, the synthetic population is just random draws from the probabilistic graphical model.
Though we can take marginal distributions of all variables as supplement for the graphical model, the population generated by Bayesian network model can only precisely match the marginals of mother nodes.
This is because the distribution of a node is uniquely defined by its parent node and the conditional distribution based on PGM.
And this point should be taken into consideration when choosing the population synthesis method.

\section{Data}
Most paper in population synthesis used Public Use Microdata Sample (PUMS) data, which is available both in individual level and household level.
It includes multiple socio-economic parameters of a household or an individual in the region.
The sample rate of this dataset is about 1\% to 10\% of the total population in the region. 
PUMS data is very complete in US cities and can be easily downloaded from government census website \footnote{https://www.census.gov/programs-surveys/acs/data/pums.html}, but the same data for cities in developing countries can be relatively outdated and harder to find. The sample data is used as seeds to initialize the contingency table in IPF methods. Another important part of data is called aggregate data \footnote{https://factfinder.census.gov}. They give marginal distribution for variables of our interests. It's worth noticing that the marginal distributions for subsets with more than 2 variables can be rarely found. An ideal format for aggregate data is that we have the marginal distributions in the level of geographical zone cause IPF is carried out separately in each zone. So when downloading the aggregate data, one should first choose a geographic type in corresponding to the zones in the virtual city, then specify the variables or tables you want.

The choice of algorithm should be based both on the data we have and the appearance of the final result we want.
What is also important for us is the simplicity and novelty of methodology. We will start from a conventional IPF method using data from Houston city, then according to the execution time and the accuracy of the result we will make further extension to IPF method.

\section{Network processing}
\subsection{OSM download}
Currently we use MapZen to download the metro area network.
We still have to isolate only relevant polygons however.

\subsection{GIS2FE}
We use this tool to convert \texttt{...\_osm\_line.shp} files to a network.
This tool was developed by \cite{karduni2016protocol,}.

@@

In the next step, the synthetic population shall be constructed by selecting households from PUMS in proportion to the estimated weights in the multiway table. Or we can do Monte Carlo sampling method to draw households from estimated weights from IPF.
Specifically, the number of households of each demographic type in a census tract can be obtained by multiplying the total number of households by the probability in the cell representing this household type, or by drawing the numbers at random according to these probabilities.
However, the household type represent by a cell in multiway table is not exactly that of PUMS data.
The fitted multiway table can only be span by control variables, which is defined as variables which have marginal distribution in census data.
However, households and individuals in disaggregate data are usually described by additional desired variables.
So, the probability of a household type in PUMS data being chosen shall be assigned according to the distance between such a household $p$ and a household type $c$ in multi-way table.
In \cite{beckman1996creating}, this probability is defined as 
\begin{equation}
	D(p,c) = w_p \prod_{i \in J}\Big(1-|(d_i^p-d_i^c)/r_i|^k\Big)*\prod_{i \notin J}\Big(1-\Delta(d_i^p, d_i^c)\Big)
\end{equation}

@@

\begin{enumerate}[Step 1:]
\item Network: modify OSM network file, make it legal input to SimMobility
  \begin{enumerate}[1.]
  \item links has to be unidirectional
  \item shorten the links and add turning points
  \item maintain the attributes of each object in the network
  \end{enumerate}
\item Population Synthesis
  \begin{enumerate}[1.]
  \item Data collection: sample data, aggregate data for zones in each city
  \item Data processing: categorize sample data and get joint distribution, formatting aggregate data
  \item Apply IPF algorithm and use the weights result to draw synthetic population
  \item Assign households in each county to locations based on land use information
  \end{enumerate} 
\end{enumerate}

@@

\subsection{Data processing and IPF}
In detail, the data processing and IPF includes following steps: Getting marginal distribution of variables that we are interested in from Census Bureau; Processing the data of marginal distributions to convenient format; Getting sample data from PUMS, and selecting sample data in our metro area by mapping census tract to county; Selecting variables in sample data, categorizing sample data based on "PUMS codebook" and variables categories in aggregate data; Generating joint distribution of sample data, which means get the frequency of each kind of combination of characteristics in sample data; Using joint distribution of sample data and marginal distribution as input to IPF; Performing IPF algorithm; Using output joint distribution of IPF to draw synthetic population.

\begin{figure}[h]
	\centering
	\includegraphics[width = 0.5\textwidth]{ipf_flow_chart.png}
	\caption{Flow chart of data processing and IPF}
	\label{fig: flowchart}
\end{figure}

In our first exploration, we have used 5 attributes to describe a households and they are household type, household size, household income in the last 12 months, number of vehicles and number of workers in the household. We not only find one dimensional marginal distribution for those variables but also two dimensional marginal distributions for household type $\times$ household size, number of workers $\times$ household size as well as number of vehicles $\times$ number of workers in the household. Across the process of data retrieving and processing, we find the categories each variable can have is largely depends on the availability and format of aggregate data.

@@

\newpage
%\section{References}
\printbibliography
\end{document}
